An algorithm for finding the global extremum of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-08-29 2023/MMST/samplepaper.tex
add some fix

\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, use formula (\ref{eq14}). In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\includegraphics[width=\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach \#3} \label{fig3}
\end{figure}

@@

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
\end{figure}

@@

\subsubsection{Description of the modification.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
  \item The value $z_v^*$ depends on the index value
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r > 0$ is a parameter of the method.

@@

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one boundary point and one non-computable point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}

@@

The numerical experiments were carried out using a computer with Intel® Core™ i7-10750H CPU @ 2.60GHz using the developed software package implementing the modification described in Section 3. To generate a series of problems, GKLS generator described in \cite{Gaviano2003} was used. It allows generating the multiextremal optimization problems with the properties known in advance. These problems were appended with the non-computable regions of various kinds.

@@

\begin{table}[h]
\caption{Comparison of the original and modified GSA}\label{tab1}
\begin{tabular}{|l|c|c|c|}
\hline
 &  classical  &  round district  &  random district  \\
\hline
Average number of trials & 781,66	& 854,91 & 839,03 \\
Average number of trials in non-computability regions & - & 185,79 & 42,31 \\
Number of solved problems & 100 / 100 & 100 / 100 & 100 / 100\\
\hline
\end{tabular}
\end{table}

\begin{figure}[h]
\includegraphics[width=\textwidth]{fig6.png}
\caption{Examples of problems \#\# 19, 25, and 35 from the first series of experiments} \label{fig6}
\end{figure}

@@

The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem \#35. At that, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})-(\ref{eq24}) that was observed in all problems presented as the examples.

@@

@@

\begin{table}
\caption{Problem series of various dimensions}\label{tab2}
\begin{tabular}{|l|l|c|c|c|}
\hline
 N &  Task type & Average num. of trials & Average num. of non-comput. & Solved problems \\
\hline
$3$ & classical & 12251.1 & - & 100 / 100 \\
$ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
$ $ & random & 12210.7 & 63.4 & 100 / 100 \\
\hline
$4$ & classical & 25144.5 & - & 100 / 100 \\
$ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
$ $ & random & 25287.5 & 26.4 & 100 / 100 \\
\hline
$5$ & classical & 56725.9 & - & 100 / 100 \\
$ $ & round & 47840.1 & 24109.2 & 100 / 100\\
$ $ & random & 57874.8 & 82.6 & 100 / 100\\
\hline
\end{tabular}
\end{table}

>>>>

\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, use formula (\ref{eq14}). In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\includegraphics[width=\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach No. 3} \label{fig3}
\end{figure}

@@

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}
\end{figure}

@@

\subsubsection{Description of the modified algorithm.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
  \item The value $z_v^*$ depends on the index value
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r > 0$ is a parameter of the method.

@@

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}

@@

The numerical experiments were carried out using a computer with Intel\textregistered\ Core\texttrademark\ i7-10750H CPU @ 2.60GHz using the developed software package implementing the modification described in Section 3. To generate a series of problems, GKLS generator described in \cite{Gaviano2003} was used. It allows generating the multiextremal optimization problems with the properties known in advance. These problems were appended with the non-computable regions of various kinds.

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{fig6.png}
\caption{Examples of problems No.No. 19, 25, and 35 from the first series of experiments} \label{fig6}
\end{figure}

\begin{table}[h]
\caption{Comparison of the original and modified GSA}\label{tab1}
\centering
\begin{tabular}{|l|c|c|c|}
\hline
 &  { classical }  & { round district }  & { random district }  \\
\hline
Average number of trials & 781.7	& 854.9 & 839.0 \\
\begin{tabular}{@{}l@{}} {Average number of trials } \\  { in non-computability regions } \end{tabular}  & - & 185.8 & 42.3 \\
{Number of solved problems } & 100 / 100 & 100 / 100  &  100 / 100 \\
\hline
\end{tabular}
\end{table}

@@

The use of rule (\ref{eq25}) forms a grid of trials inside the non-computability regions that is visible clearly for Problem No. 35. At that, the trial points are denser at the boundary of the non-computable region close to potential minimum when attempting to approach this one from different sides due to the computational rules (\ref{eq22})-(\ref{eq24}) that was observed in all problems presented as the examples.

@@

\begin{table}
\caption{Problem series of various dimensions}\label{tab2}
\centering
\begin{tabular}{|l|l|c|c|c|}
\hline
 N &  Task type & Average number of trials & \begin{tabular}{@{}c@{}} {Average number of trials} \\  {in non-computability regions} \end{tabular} & Solved problems \\
\hline
$3$ & classical & 12251.1 & - & 100 / 100 \\
$ $ & round & 12201.8 & 4763.6 & 100 / 100 \\
$ $ & random & 12210.7 & 63.4 & 100 / 100 \\
\hline
$4$ & classical & 25144.5 & - & 100 / 100 \\
$ $ & round & 29840.3 & 15004.4  & 100 / 100 \\
$ $ & random & 25287.5 & 26.4 & 100 / 100 \\
\hline
$5$ & classical & 56725.9 & - & 100 / 100 \\
$ $ & round & 47840.1 & 24109.2 & 100 / 100\\
$ $ & random & 57874.8 & 82.6 & 100 / 100\\
\hline
\end{tabular}
\end{table}

@@
